---
title: "Google Data Analytics"
author: "Meet"
date: "2023-04-02"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

### Importing Library

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(chron) # For time 
```

## Importing Data

```{r}
march_22 <- read.csv("Biker_03_2022.csv")
april_22 <- read.csv("Biker_04_2022.csv")
may_22 <- read.csv("Biker_05_2022.csv")
june_22 <- read.csv("Biker_06_2022.csv")
july_22 <- read.csv("Biker_07_2022.csv")
august_22 <- read.csv("Biker_08_2022.csv")
september_22 <- read.csv("Biker_09_2022.csv")
october_22 <- read.csv("Biker_10_2022.csv")
november_22 <- read.csv("Biker_11_2022.csv")
december_22 <- read.csv("Biker_12_2022.csv")
january_23 <- read.csv("Biker_01_2023.csv")
february_23 <- read.csv("Biker_02_2023.csv")
```

Combine to one data set named as *alldata*.

```{r}
alldata <- rbind(march_22, april_22, may_22, june_22, july_22, august_22, september_22, october_22, november_22, december_22, january_23, february_23)
```

Get a overview of the dataset.

```{r}
head(alldata)
glimpse(alldata)
summary(alldata)
```

## Cleaning data

To avoid the future complications, let's first make the missing and spacing values if any into **NA**

```{r}
alldata[alldata=="" | alldata==" "] <- NA
```

### Check NA values

Let's check how many missing values are there in the dataset column-wise.

#### ride_id

```{r}
table(is.na(alldata$ride_id))
```

#### rideable_type

```{r}
table(is.na(alldata$rideable_type))
```

#### started_at

```{r}
table(is.na(alldata$started_at))
```

#### ended_at

```{r}
table(is.na(alldata$ended_at))
```

#### start_station_name

```{r}
table(is.na(alldata$start_station_name))
```

Now, we know that there are some missing value *(850418)* in this column so if needed we can add...

#### start_station_id

```{r}
table(is.na(alldata$start_station_id))
```

This column have *850550* missing values

#### end_station_name

```{r}
table(is.na(alldata$end_station_name))
```

This column have *909038* missing values

#### end_station_id

```{r}
table(is.na(alldata$end_station_id))
```

This column have *909179* missing values

#### start_lat

```{r}
table(is.na(alldata$start_lat))
```

#### start_lng

```{r}
table(is.na(alldata$start_lng))
```

#### end_lat

```{r}
table(is.na(alldata$end_lat))
```

This column have *5938* missing values

#### end_lng

```{r}
table(is.na(alldata$end_lng))
```

This column have *5938* missing values

#### member_casual

```{r}
table(is.na(alldata$member_casual))
```

#### ride_time

```{r}
table(is.na(alldata$ride_time))
```

#### day_of_week

```{r}
table(is.na(alldata$day_of_week))
```

## Analysis

Let's plot the counts of bike rides on each day of the week with different types of riders

```{r}
ggplot(data=alldata) +
  geom_bar(mapping=aes(x=day_of_week, fill=rideable_type)) +
  facet_wrap(~member_casual) +
  scale_y_continuous(labels = scales::comma) +
  annotate("text",x=1, y=150000, label="Sunday", angle=90)+
  annotate("text",x=2, y=150000, label="Monday", angle=90)+
  annotate("text",x=3, y=150000, label="Tuesday", angle=90)+
  annotate("text",x=4, y=150000, label="Wednesday", angle=90)+
  annotate("text",x=5, y=150000, label="Thursday", angle=90)+
  annotate("text",x=6, y=150000, label="Friday", angle=90)+
  annotate("text",x=7, y=150000, label="Saturday", angle=90) +
  labs(title="No. of Rides with different types of bikes", x="Day of the week", y="Counts")
ggsave("No. of Rides with different types of bikes.png")
```

### Manipulating Data

Add some important column which will be useful for our analysis to the *alldata* dataset like *date*, *month*, *day*, *year*

```{r}
date <- c(1:5829084)
month <- c(1:5829084)
day <- c(1:5829084)
year <- c(1:5829084)
df <- data.frame(date,month,day,year)
df$date[date!=""] <- NA
df$month[month!=""] <- NA
df$day[day!=""] <- NA
df$year[year!=""] <- NA
```

```{r}
alldata <- cbind(alldata, df)
```

As *started_at* is in the "character" type so lets first convert it to "factor(integer)" form for further making it separate *DateTime* format to *Date*

```{r}
alldata$started_at <- factor(alldata$started_at)
```

```{r}
typeof(alldata$started_at)
```

Now, convert "factor" to "DateTime" format

```{r}
alldata$started_at <- mdy_hm(alldata$started_at)
```

and finally separate the Date out

```{r}
alldata$date <- as.Date(alldata$started_at)
```

Now separate year, month and day

```{r}
alldata$year <- year(alldata$started_at)
```

```{r}
alldata$month <- month(alldata$started_at)
```

```{r}
alldata$day <- day(alldata$started_at)
```

### Plotting Data

Let's first separate out a dataset in which there will be data of membership type, day, and count of number of rides in a grouped form...

```{r}
rides_data <- alldata %>% 
  group_by(member_casual, day_of_week) %>% 
  summarize(number_of_rides=n()) %>% 
  select(member_casual, day_of_week, number_of_rides)
```

Plotting the data

```{r}
ggplot(data=rides_data) +
  geom_col(mapping=aes(x=day_of_week, y=number_of_rides), fill="pink") +
  facet_wrap(~member_casual) + 
  scale_y_continuous(labels = scales::comma) +
  annotate("text",x=1, y=150000, label="Sunday", angle=90)+
  annotate("text",x=2, y=150000, label="Monday", angle=90)+
  annotate("text",x=3, y=150000, label="Tuesday", angle=90)+
  annotate("text",x=4, y=150000, label="Wednesday", angle=90)+
  annotate("text",x=5, y=150000, label="Thursday", angle=90)+
  annotate("text",x=6, y=150000, label="Friday", angle=90)+
  annotate("text",x=7, y=150000, label="Saturday", angle=90)+
  labs(title="Total rides on each day of the week", x="Day of the week", y="No. of rides")
ggsave("Total rides on each day of the week.png")
```

#### Manipulating Data

Let's find the total ride time to gain some insights\
First convert *ride_time* from "character" to "integer" in *alldata* dataset

```{r}
alldata$ride_time <- factor(alldata$ride_time)
```

Add one column *average_ride_time* into our newly created dataset

```{r}
average_ride_time <- c(1:14)
average_ride_time <- data.frame(average_ride_time)
average_ride_time[!is.na(average_ride_time)] <- NA
rides_data <- cbind(rides_data, average_ride_time)
```

Manually enter the values using this code to find the mean of each row in our new dataset, **rides_data**

```{r}
mean(times((alldata %>% filter(member_casual=="member", day_of_week==7))$ride_time))
```

Fill the data found by the above code

```{r}
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==1] <- "00:24:49"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==2] <- "00:22:02"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==3] <- "00:19:17"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==4] <- "00:18:38"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==5] <- "00:19:20"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==6] <- "00:20:20"
rides_data$average_ride_time[rides_data$member_casual=="casual"&rides_data$day_of_week==7] <- "00:24:24"
```

```{r}
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==1] <- "00:13:36"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==2] <- "00:11:50"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==3] <- "00:11:37"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==4] <- "00:11:42"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==5] <- "00:11:52"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==6] <- "00:12:05"
rides_data$average_ride_time[rides_data$member_casual=="member"&rides_data$day_of_week==7] <- "00:13:43"
```

```{r}
typeof(rides_data$average_ride_time)
```

Values filled is of "character" format, converting into "factor"

```{r}
rides_data$average_ride_time <- factor(rides_data$average_ride_time)
```

```{r}
summary(rides_data)
```

Just verifying...

```{r}
sum(rides_data$number_of_rides)
```

Making data more readable by making some minor upgrades

```{r}
rides_data$day_of_week[rides_data$day_of_week==1] <- "Sunday"
rides_data$day_of_week[rides_data$day_of_week==2] <- "Monday"
rides_data$day_of_week[rides_data$day_of_week==3] <- "Tuesday"
rides_data$day_of_week[rides_data$day_of_week==4] <- "Wednesday"
rides_data$day_of_week[rides_data$day_of_week==5] <- "Thursday"
rides_data$day_of_week[rides_data$day_of_week==6] <- "Friday"
rides_data$day_of_week[rides_data$day_of_week==7] <- "Saturday"
```

#### Export New Dataset

```{r}
write.csv(rides_data, "Final Analysis.csv")
```

Create another dataset to analysis number of rides month-wise

```{r}
rides_data_monthly <- alldata %>% 
  group_by(member_casual, month) %>% 
  summarize(number_of_rides=n()) %>% 
  select(member_casual, month, number_of_rides)
```

Plot the data...

```{r}
ggplot(data=rides_data_monthly) +
  geom_col(mapping=aes(x=month, y=number_of_rides), fill="orange")+
  facet_wrap(~member_casual)+
  scale_y_continuous(labels = scales::comma)+
  labs(title = "No. of rides Month-wise", x="Month", y="No. of Rides")+
  annotate("text",x=1, y=100000, label="January", angle=90)+
  annotate("text",x=2, y=100000, label="February", angle=90)+
  annotate("text",x=3, y=100000, label="March", angle=90)+
  annotate("text",x=4, y=100000, label="April", angle=90)+
  annotate("text",x=5, y=100000, label="May", angle=90)+
  annotate("text",x=6, y=100000, label="June", angle=90)+
  annotate("text",x=7, y=100000, label="July", angle=90)+
  annotate("text",x=8, y=100000, label="August", angle=90)+
  annotate("text",x=9, y=100000, label="September", angle=90)+
  annotate("text",x=10, y=100000, label="October", angle=90)+
  annotate("text",x=11, y=100000, label="November", angle=90)+
  annotate("text",x=12, y=100000, label="December", angle=90)
ggsave("No. of rides Month-wise.png")
```

Make data more readable and export it for further possible use

```{r}
rides_data_monthly$month[rides_data_monthly$month==1] <- "January"
rides_data_monthly$month[rides_data_monthly$month==2] <- "February"
rides_data_monthly$month[rides_data_monthly$month==3] <- "March"
rides_data_monthly$month[rides_data_monthly$month==4] <- "April"
rides_data_monthly$month[rides_data_monthly$month==5] <- "May"
rides_data_monthly$month[rides_data_monthly$month==6] <- "June"
rides_data_monthly$month[rides_data_monthly$month==7] <- "July"
rides_data_monthly$month[rides_data_monthly$month==8] <- "August"
rides_data_monthly$month[rides_data_monthly$month==9] <- "September"
rides_data_monthly$month[rides_data_monthly$month==10] <- "October"
rides_data_monthly$month[rides_data_monthly$month==11] <- "November"
rides_data_monthly$month[rides_data_monthly$month==12] <- "December"
```

```{r}
write.csv(rides_data_monthly, "Monthly Analysis.csv")
```

Now after applying formula in the Excel, update the dataset to add the hour of rides by every row in *rides_data* dataset

```{r}
average_time_analysis <- read.csv("Average Time Analysis in Hour.csv")
```

```{r}
head(average_time_analysis)
summary(average_time_analysis)
```

```{r}
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Sunday"] <- 1
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Monday"] <- 2
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Tuesday"] <- 3
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Wednesday"] <- 4
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Thursday"] <- 5
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Friday"] <- 6
average_time_analysis$day_of_week[average_time_analysis$day_of_week=="Saturday"] <- 7
```

Plotting data to analyze ride hours and days of the week relationship

```{r}
ggplot(data=average_time_analysis) +
  geom_col(mapping=aes(x=day_of_week, y=in_hour), fill="green")+
  facet_wrap(~member_casual) + 
  scale_y_continuous(labels = scales::comma) +
  annotate("text",x=1, y=35000, label="Sunday", angle=90)+
  annotate("text",x=2, y=35000, label="Monday", angle=90)+
  annotate("text",x=3, y=35000, label="Tuesday", angle=90)+
  annotate("text",x=4, y=35000, label="Wednesday", angle=90)+
  annotate("text",x=5, y=35000, label="Thursday", angle=90)+
  annotate("text",x=6, y=35000, label="Friday", angle=90)+
  annotate("text",x=7, y=35000, label="Saturday", angle=90)+
  labs(title="Total time(in Hours) on each day of the week", x="Day of the week", y="Total hours of Ride")
ggsave("Total time(in Hours) on each day of the week.png")
```

Based on this Analysis, we will make a presentation regarding our business goal\
Thank you...
